### 2014-01-15 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
More work on main text of dpp manuscript.

Following the recommendations of \citet{Oaks2012}, we modify the \msb model to
test whether alternative parameterizations and priors will improve the behavior
of the approximate-Bayesian model-choice approach.
We implement a Dirichlet-process prior (DPP) over all possible models of
divergence, and use gamma and beta probability distributions in place of
uniform priors on many of the model's parameters.
Our simulation-based analyses show that the new implementation has improved
robustness, accuracy, and power compared to the original model.
Also, our results show that both the uniform priors on demographic and
divergence-time parameters and the U-shaped prior on divergence models
contribute to the biases of \msb.
When the Oaks et al.'s \citeyear{Oaks2012} sequence data of 22 pairs of taxa
from the Philippines is reanalyzed under the new model, we find a large amount
of posterior uncertainty in the number of divergence events shared among the
taxa; a result that congruent with intuition given the richness of the model
and the relatively small amount of information in the data.

@@

Overall, the result suggest that the \modelDPP and \modelUniform models are
relatively robust in terms of model-choice accuracy, and when model violations
do cause them to be biased, they tend to under-estimate the probability of the
biogeographically interesting model of a single divergence event.
In contrast, the \modelOld model is very sensitive to model violations, and
strongly over-estimates the one-divergence model whenever the model is
misspecified.
Furthermore, the results suggest that using exponentially distributed priors on
nuisance parameters rather than uniform priors helps the \modelUshaped model
perform better than \modelOld, but it is still hindered by the U-shaped prior
on divergence models and tends to over-estimate the probability of the
one-divergence model whenever there are violations of the model.

@@

\subsection*{Power analyses: Model choice}
Our modifications of the \msb model decrease the method's bias toward clustered
divergences when applied to data generated under random divergence times
(Figure~\ref{figPowerPsiOld4} \& 
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelOld model performs the worst of the four models across
all three series of data-generating models, inferring a single divergence event across
most of the 18,000 simulations
(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Importantly, the \modelOld tends to strongly support these estimates
of one divergence across most of the simulations
(Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
The \modelDPP model also prefers the one-divergence model 
when divergences are random over narrow windows of time, but
performs much better when divergences are random over a timescale
of 1--2 coalescent units
(Figure~\ref{figPowerPsiOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
However, even when \modelDPP infers the one-divergence model over narrow
timescales, the posterior probability support is always low
(Figure~\ref{figPowerPsiProbOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
% Even when the simulated data are identically distributed as the \modelOld model
% (except for the divergence times) the \modelOld is still strongly biased
% towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
% model is the least likely to infer a single event when applied to these data
% (Figure \ref{figPowerPsiOldDPP}).
The \modelUniform model never infers the one-divergence model in any of the
simulation replicates but still tends to infer relatively few (4--6) divergence
events when divergences are random over longer periods 
(Figure~\ref{figPowerPsiOld4} I--L \& plots M--R of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Using exponential priors on divergence-time and theta parameters does increase
the power of the \modelUshaped model compared to \modelOld across all three
series of data models, but the U-shaped prior still prevents the model from
performing as well as the \modelDPP and \modelUniform models 
(Figure~\ref{figPowerPsiOld4} \& 
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).

The improved power of the new models is even more pronounced when looking at
estimates of the variance of divergence times (\divTimeDispersion) across the
simulations
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
The performance among the models is so different, that the histograms of
\divTimeDispersion estimates cannot be plotted along a shared x-axis.
The \modelDPP and \modelUniform models perform similarly across all three
series of data models, inferring values of \divTimeDispersion consistent with
one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
replicates across all the simulations.
In contrast, the \modelOld model infers values consistent with a single divergence event in
most of the replicates across all the simulations.
Using exponential priors on divergence-time and theta parameters greatly
increases the power of the \modelUshaped model to detect variation in
divergence times relative to \modelOld, but it still has less power than the
models with Dirichlet-process or uniform priors across divergence models 
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).

As mentioned above, we also see increased power of the new models when
looking at the estimated posterior probability of the one-divergence model
across the power analyses 
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models estimate low posterior probability of
$\divTimeNum = 1$ across all of the simulations.
This is in contrast to the \modelOld model, which infers high posterior probabilities of a single
divergence for most replicates across all simulations
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The exponential priors on divergence-time and theta parameters (model
\modelUshaped) result in lower estimates of the probability of one divergence
when compared to \modelOld, but higher estimates when compared to \modelUniform
and \modelDPP
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models do frequently support the one-model
according to a Bayes factor criterion of greater than 10, however still less
frequently than the \modelOld model.
This result is not surprising given the extremely small prior probability of
the one-divergence model under the \modelDPP and \modelUniform models (I.e.,
very few posterior samples of the one-divergence model will result in a
large Bayes factor under these models).
However, the small posterior probability of the one-divergence model estimated
under \modelDPP and \modelUniform should prevent an investigator from
overinterpreting the Bayes factor as strong support for clustered divergences.

@@

\subsection*{Empirical results}
As expected based on the results of \citet{Oaks2012}, when the Philippines data
are analyzed under the \empModelOld model, there is strong support for very few
divergence events shared among all 22 pairs of taxa
(Figure~\ref{figPhilippines}A).
When these data are analyzed under models allowed under the new implementation,
there is much less support for highly clustered models and much greater
uncertainty regarding the number of divergence events shared among the taxa,
especially under the DPP models (Figure~\ref{figPhilippines}B--E).
Figure \ref{figPhilippines} also shows the prior distribution across the number
of divergence events (\divTimeNum) for each of the model, as well as the
average prior probability of an unordered and ordered model of divergence
(\divTimeIndexVector) across \divTimeNum for each model.
Estimates under the new models tend to be similar to the prior, which is
expected under such a parameter-rich model when there is limited information
from the data (four summary statistics from a single locus for each pair of
taxa).

The disparity between the results of the \empModelOld model and the new models
is even more pronounced when we look at the 10 divergence models
(\divTimeMapVector) estimated to have the highest probability under each of the
models
(Figures~S\labelcref{figDivModelsDPP,figDivModelsDPPInform,figDivModelsDPPSimple,figDivModelsUniform,figDivModelsOld}).
Again, the new models estimate more divergences, a large amount of posterior
uncertainty, and an order of magnitude smaller probability for the top model
when compared to the \empModelOld model.

Figure \ref{figNegrosPanay} shows the estimated posterior probability
distribution over the number of divergence events when the data from the
nine-taxon pairs of the Islands of Negros and Panay are analyzed under the DPP
models that sample over unordered (\npModelDPP) and ordered
(\npModelDPPOrdered) models of divergence.
The result are similar under both models and, again, yield a large amount of
uncertainty about the number of divergence events that is similar to the prior
uncertainty.
Both models estimate a similar set of 10 unordered divergence models with the
highest posterior probability
(Figures~S\labelcref{figDivModelsNP,figDivModelsNPOrdered}).

@@

\section*{Discussion}
Our results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In our implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to model violations and, when present,
strongly over-estimates the probability of one divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
When more appropriate priors are used for divergence-time and theta parameters,
and either a Dirichlet-process or uniform prior applied across divergence
models, the model is less sensitive to violations, and, when they do cause
bias, the method is conservative and tends to under-estimate the probability of
biogeographically interesting models of temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).

Furthermore, the modifications to the model also improve the method's power to
detect random variation in divergence times, reducing the bias toward
estimating clustered divergences
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
Our results are similar to those of \citet{Oaks2012} in that we also find
the \msb model to infer strong support for clustered divergences when
divergences are random over quite broad timescales
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}.
Our results expand on this by showing that this behavior is consistent
across a range of conditions underlying the data.
When using the new priors, the method is less prone to spurious inference of
clustered divergence models, and much less likely to incorrectly infer such
models with strong support.

By evaluating a model intermediate between the old and new implementation
(\modelUshaped), we were able to determine the relative effects of the
modifications to the model.
Across all of our analyses, the results show that using better priors on
divergence-time and nuisance parameters alone does improve the performance of
the method.
The magnitude of the bias toward inferring support for the one-divergence
model when there are model violations is reduced when the exponential
priors are used in place of the uniform priors
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
Furthermore, better priors on nuisance parameters improves the method's
power to detect temporally random divergences.
Throughout our analyses, the intermediate model (\modelUshaped) performs
better than the \msb model, but not as well as the models with alternative
priors on divergence models.
This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
models of temporally clustered divergences is caused by a combination of
(1) uniform priors on parameters hindering the marginal likelihoods of models
with more \divTime{} parameters and
(2) the U-shaped prior on divergence models giving low prior density to models
with intermediate numbers of divergence parameters.

While our modifications improve the behavior of the model, we urge
caution when using the method and interpreting its results.
The method attempts to approximate the posterior of a very parameter-rich model
using relatively little information from the data.
For example, when applied to the dataset of Philippine vertebrates analyzed by
\citet{Oaks2012} the model can have as many as 604--625 free parameters
(depending on \divTimeNum), and samples over 1002 divergence models.
Our modifications allow this model to be simplified to 471--492 free
parameters.
Sampling over hundreds of parameters and more than a thousand models is a
difficult inference problem, especially when only using a small number of
summary statistics calculated from the sequence alignments of each taxon pair.
The population-genetic summary statistics used by our method contain little
information about many of the free parameters in the model.
Thus, we expect the improved method will still be sensitive to priors, and the
power, while improved, may still be low.
Furthermore, ABC methods in general have been shown to be biased
\citep{Robert2011} for model choice.
As recommended by \citet{Oaks2012}, any results from the method should be
accompanied by assessments of prior sensitivity and power.
We encourage investigators to view this method and its predecessor as a means
of exploring their data rather than a rigorous means of evaluating hypotheses.

Given the difficulty of this estimation problem, we anticipate that
full-likelihood methods that can leverage all of the information present in the
sequence data will become increasingly important for robustly estimating the
temporal history of divergences across taxa \citep{JeetDiss}.
With improving numerical methods for sampling over models of differing
dimensionality \citep{Green1995,Lemey2009} and increasing efficiency of
likelihood calculations \citep{Ayres2012}, analyzing rich comparative
phylogeograpical models in a full-likelihood Bayesian framework is becoming
computationally practical, especially when considering that simulating millions
of random datasets from the prior under the simple ABC rejection approach is
inefficient and computationally nontrivial.

>>>>

Following the recommendations of \citet{Oaks2012}, we modify the \msb model to
test whether alternative parameterizations and priors will improve the behavior
of the approximate-Bayesian model-choice approach.
We implement a Dirichlet-process prior (DPP) over all possible models of
divergence, and use gamma and beta probability distributions in place of
uniform priors on many of the model's parameters.
Our simulation-based analyses show that the new implementation has improved
robustness, accuracy, and power compared to the original model.
Also, our results show that both the uniform priors on demographic and
divergence-time parameters and the U-shaped prior on divergence models
contribute to the biases of \msb.
When the Oaks et al.'s \citeyear{Oaks2012} sequence data of 22 pairs of taxa
from the Philippines is reanalyzed under the new model, we find a large amount
of posterior uncertainty in the number of divergence events shared among the
taxa; a result congruent with intuition given the richness of the model
and the relatively small amount of information in the data.

@@

Overall, the results suggest that the \modelDPP and \modelUniform models are
relatively robust in terms of model-choice accuracy, and when model violations
do cause them to be biased, they tend to under-estimate the probability of the
biogeographically interesting model of a single divergence event.
In contrast, the \modelOld model is very sensitive to model violations, and
strongly over-estimates the one-divergence model whenever the model is
misspecified.
Furthermore, the results suggest that using exponentially distributed priors on
nuisance parameters rather than uniform priors helps the \modelUshaped model
perform better than \modelOld, but it is still hindered by the U-shaped prior
on divergence models and tends to over-estimate the probability of the
one-divergence model whenever there are violations of the model.

@@

\subsection*{Power analyses: Model choice}
Our modifications of the \msb model decrease the method's bias toward clustered
divergences when applied to data generated under random divergence times
(Figure~\ref{figPowerPsiOld4} \& 
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelOld model performs the worst of the four models across
all three series of data-generating models, inferring a single divergence event across
most of the 18,000 simulations
(Figure~\ref{figPowerPsiOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Importantly, the \modelOld tends to strongly support these estimates
of one divergence across most of the simulations
(Figure~\ref{figPowerPsiProbOld4}A--D \& plots A--F of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
The \modelDPP model also prefers the one-divergence model 
when divergences are random over narrow windows of time, but
performs much better when divergences are random over a timescale
of 1--2 coalescent units
(Figure~\ref{figPowerPsiOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
However, even when \modelDPP infers the one-divergence model over narrow
timescales, the posterior probability support is always low
(Figure~\ref{figPowerPsiProbOld4}M--P \& plots S--X of Figures
S\labelcref{figPowerPsiProbOld,figPowerPsiProbUniform,figPowerPsiProbExp}).
% Even when the simulated data are identically distributed as the \modelOld model
% (except for the divergence times) the \modelOld is still strongly biased
% towards a single event (Figure \ref{figPowerPsiOldOld}), whereas the \modelDPP
% model is the least likely to infer a single event when applied to these data
% (Figure \ref{figPowerPsiOldDPP}).
The \modelUniform model never infers the one-divergence model in any of the
simulation replicates but still tends to infer relatively few (4--6) divergence
events when divergences are random over longer periods 
(Figure~\ref{figPowerPsiOld4} I--L \& plots M--R of Figures
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).
Using exponential priors on divergence-time and \myTheta{} parameters does
increase the power of the \modelUshaped model compared to \modelOld across all
three series of data models, but the U-shaped prior still prevents the model
from performing as well as the \modelDPP and \modelUniform models 
(Figure~\ref{figPowerPsiOld4} \& 
S\labelcref{figPowerPsiOld,figPowerPsiUniform,figPowerPsiExp}).

The improved power of the new models is even more pronounced when looking at
estimates of the variance of divergence times (\divTimeDispersion) across the
simulations
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
The performance among the models is so different, that the histograms of
\divTimeDispersion estimates cannot be plotted along a shared x-axis.
The \modelDPP and \modelUniform models perform similarly across all three
series of data models, inferring values of \divTimeDispersion consistent with
one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
replicates across all the simulations.
In contrast, the \modelOld model infers values consistent with a single
divergence event in most of the replicates across all the simulations.
Using exponential priors on divergence-time and \myTheta{} parameters greatly
increases the power of the \modelUshaped model to detect variation in
divergence times relative to \modelOld, but it still has less power than the
models with Dirichlet-process or uniform priors across divergence models 
(Figure~\ref{figPowerOmegaOld4} \& 
S\labelcref{figPowerOmegaOld,figPowerOmegaUniform,figPowerOmegaExp}).
An estimated value of \divTimeDispersion less than 0.01 is an arbitrary
threshold for ``simultaneous'' divergence used throughout the \msb
literature.
Albeit arbitrary, \cite{Oaks2012} did show via simulation that the true value
of \divTimeDispersion will almost always be greater than 0.01 when divergences
are random over periods of 0.1 \globalcoalunit generations or more (see
Figure~S4 of \citet{Oaks2012}).

As mentioned above, we also see increased power of the new models when
looking at the estimated posterior probability of the one-divergence model
across the power analyses 
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models estimate low posterior probability of
$\divTimeNum = 1$ across all of the simulations.
This is in contrast to the \modelOld model, which infers high posterior probabilities of a single
divergence for most replicates across all simulations
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The exponential priors on divergence-time and \myTheta{} parameters (model
\modelUshaped) result in lower estimates of the probability of one divergence
when compared to \modelOld, but higher estimates when compared to \modelUniform
and \modelDPP
(Figure~\ref{figPowerPsiProbOld4} \& 
S\labelcref{figPowerPsiProbOld,figPowerPsiUniform,figPowerPsiExp}).
The \modelDPP and \modelUniform models do frequently support the one-divergence
model according to a Bayes factor criterion of greater than 10, however still
less frequently than the \modelOld model.
This result is not surprising given the extremely small prior probability of
the one-divergence model under the \modelDPP and \modelUniform models (I.e.,
very few posterior samples of the one-divergence model will result in a
large Bayes factor under these models).
However, the small posterior probability of the one-divergence model estimated
under \modelDPP and \modelUniform should prevent an investigator from
overinterpreting the Bayes factor as strong support for clustered divergences.

@@

\subsection*{Empirical results}
As expected based on the results of \citet{Oaks2012}, when the Philippines data
are analyzed under the \empModelOld model, there is strong support for very few
divergence events shared among all 22 pairs of taxa with a maximum \emph{a
posteriori} (MAP) estimate of one-shared divergence
(Figure~\ref{figPhilippines}A).
When these data are analyzed under models allowed under the new implementation,
there is much less support for highly clustered models and much greater
uncertainty regarding the number of divergence events shared among the taxa,
especially under the DPP models (Figure~\ref{figPhilippines}B--E).
Figure \ref{figPhilippines} also shows the prior distribution across the number
of divergence events (\divTimeNum) for each of the model, as well as the
average prior probability of an unordered and ordered model of divergence
(\divTimeIndexVector) across \divTimeNum for each model.
Estimates under the new models tend to be similar to the prior, which is
expected under such a parameter-rich model when there is limited information
from the data (four summary statistics from a single locus for each pair of
taxa).

The disparity between the results of the \empModelOld model and the new models
is even more pronounced when we look at the 10 divergence models
(\divTimeIndexVector) estimated to have the highest probability under each of the
models
(Figures~S\labelcref{figDivModelsDPP,figDivModelsDPPInform,figDivModelsDPPSimple,figDivModelsUniform,figDivModelsOld}).
Again, the new models estimate more divergences, a large amount of posterior
uncertainty, and an order of magnitude smaller probability for their respective
MAP model when compared to the \empModelOld model.

Figure \ref{figNegrosPanay} shows the estimated posterior probability
distribution over the number of divergence events when the data from the
nine-taxon pairs from the Islands of Negros and Panay are analyzed under the DPP
models that sample over unordered (\npModelDPP) and ordered
(\npModelDPPOrdered) models of divergence.
The result are similar under both models and, again, yield a large amount of
uncertainty about the number of divergence events that is similar to the prior
uncertainty.
Both models estimate a similar set of 10 unordered divergence models with the
highest posterior probability
(Figures~S\labelcref{figDivModelsNP,figDivModelsNPOrdered}).

@@

\section*{Discussion}
Our results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In our implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to model violations and, when present,
strongly over-estimates the probability of one-divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
When more appropriate priors are used for divergence-time and demographic
parameters, and either a Dirichlet-process or uniform prior applied across
divergence models, the model is less sensitive to violations, and, when they do
cause bias, the method is conservative and tends to underestimate the
probability of biogeographically interesting models of temporally clustered
divergences (Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).

Furthermore, the modifications to the model also improve the method's power to
detect random variation in divergence times, reducing the tendency to estimate
clustered divergences
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
Our results are similar to those of \citet{Oaks2012} in that we also find
the \msb model to infer strong support for clustered divergences when
divergences are random over quite broad timescales
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
Our results expand on this by showing that this behavior is consistent
across a range of conditions underlying the data.
The new method \dppmsbayes has greater power to detect random temporal variation in
divergences, is less prone to spurious inference of clustered divergence
models, and much less likely to incorrectly infer such models with strong
support
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).

By evaluating a model intermediate between the old and new implementation
(\modelUshaped), we were able to determine the relative effects of our
modifications to the model.
Across all of our analyses, the results show that using better priors on
divergence-time and demographic parameters alone does improve the performance
of the method.
The magnitude of the bias toward inferring support for the one-divergence
model when there are model violations is reduced when the exponential
priors are used in place of the uniform priors
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
Furthermore, better priors on parameters improves the method's power to detect
temporally random divergences
(Figures~\labelcref{figPowerPsiOld4,figPowerPsiProbOld4,figPowerOmegaOld4,figPowerOmegaProbOld4}).
Throughout our analyses, the intermediate model (\modelUshaped) performs better
than the \msb model, but not as well as the models with alternative priors on
divergence models.
This suggests, as predicted by \citet{Oaks2012}, that the bias of \msb toward
models of temporally clustered divergences is caused by a combination of
(1) small marginal likelihoods of models with more \divTime{} parameters due to
uniform priors on divergence-time and demographic parameters and
(2) the U-shaped prior on divergence models giving low prior density to models
with intermediate numbers of divergence parameters.
The former essentially rules out models with many \divTime{} parameters, which
causes the latter to act like an "L-shaped" prior with a spike of
prior density on the one-divergence model.
Given the parameter richness of the model and the relatively small amount of
information in the summary statistics, it is not surprising that the
combination of these two factors can create a strong tendency to infer
clustered models of divergence.

While our modifications improve the behavior of the model, we urge caution when
using the method and interpreting its results.
The method attempts to approximate the posterior of a very parameter-rich model
using relatively little information from the data.
For example, when applied to the dataset of 22 taxon pairs from the Philippines
\citet{Oaks2012} the model can have as many as 604--625 free parameters
(depending on \divTimeNum), and samples over 1002 unordered divergence models.
Even under the simplest possible model allowed under our new implementation,
the model still has 471--492 free parameters.
Furthermore, the coalescent and mutational models specified by these parameters
predict a large amount of variation in possible datasets even when the
parameter values are known.
Sampling over hundreds of parameters and more than a thousand models is a
difficult inference problem, especially when only using a small number of
summary statistics calculated from the sequence alignments of each taxon pair.
The population-genetic summary statistics used by our method contain little
information about many of the free parameters in the model.
Thus, we expect the improved method will still be sensitive to priors, and the
power, while improved, may still be low.
While we see much less prior sensitivity under the new model compared to those
observed by \citet{Oaks2012}, we still some evidence of such sensitivity when
comparing the results of the empirical data analyzed under a diffuse
(\empModelDPP) and informative (\empModelDPPInform) divergence-time prior
(Figure~\ref{figPhilippines} C versus D).
The fact that the posterior shifts toward the prior under the informative prior
suggests that the shift away from the prior toward fewer divergence events
under the diffuse prior might be still be caused by small marginal likelihoods
of models with more divergence-time parameters (Figure~\ref{figPhilippines}).

Nonetheless, it is reassuring to see a large amount of posterior uncertainty
when the new implementation is applied to the empirical datasets
(Figure~\labelcref{figPhilippines,figNegrosPanay}).
Applications of the \msb model often result in strong posterior support for
estimated scenarios
\citep{Barber2010,Bell2012,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012},
as we found here (Figure~\ref{figPhilippines}).
Given the richness of the model, the variance of the processes being modeled,
and the relatively small amount of information in the summary statistics
calculated from the sequence data, we do not expect to find strong posterior
support for any scenario.
Based on results of the empirical and power analyses
(Figures~\labelcref{figPowerPsiProbOld4,figPowerOmegaProbOld4,figPhilippines,figNegrosPanay})
the new implementation more accurately reflects posterior uncertainty and
avoids spurious support for biogeographic scenarios.

We also urge caution when applying \dppmsbayes due to the fact that ABC methods
in general are known to be biased for model choice \citep{Robert2011}.
As recommended by \citet{Oaks2012}, any results from the method should be
accompanied by assessments of prior sensitivity and power.
We encourage investigators to view this method as a means of exploring their
data rather than a rigorous means of evaluating hypotheses.

Given the difficulty of this estimation problem, we anticipate that
full-likelihood methods that can leverage all of the information present in the
sequence data will become increasingly important for robustly estimating shared
evolutionary history across taxa \citep{JeetDiss}.
With improving numerical methods for sampling over models of differing
dimensionality \citep{Green1995,Lemey2009}, advances in Monte Carlo techniques
\citep{Jordan2012}, and increasing efficiency of likelihood calculations
\citep{Ayres2012}, analyzing rich comparative phylogeograpical models in a
full-likelihood Bayesian framework is becoming computationally practical,
especially when considering that simulating millions of random datasets from
the prior under the simple ABC rejection approach is inefficient and
computationally nontrivial.
